---
title: "California Water Rate Survey Results 2017"
output:
  html_document:
    keep_md: true
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(scales)
library(raster)
library(RateParser)
library(yaml)
library(purrr)
library(fuzzywuzzyR)
library(lubridate)
library(captioner)
```

# Introduction

The [California Data Collaborative ("CaDC")](http://californiadatacollaborative.org/) is a coalition of water utilities that have pioneered a new data infrastructure non-profit 501 (c) (3) to support water managers in meeting their reliability objectives and serve the public good.

One important contribution of the CaDC was to establish a standard format and to provide the infrastructure for storage and maintainance of an open database for water rates, facilitating the work of analysts, economists and software developers interested in analyzing and understanding the differences in water rate structures and prices across many different agencies and locations. The water rate structures were organized in [Open-Water-Rate-Specification (OWRS)](https://github.com/California-Data-Collaborative/Open-Water-Rate-Specification) files, a format based on [YAML](http://yaml.org/), which is designed to be easy to store, transmit, and parse in any programming language while also being easy for humans to read.

This report presents a summary of the types of analyses and insights that can be obtained from analyzing the OWRS, especially when this information is combined with the water consumption data from water agencies and Census Data.

# Data

This report provides the combined analysis of data from 4 different sources:

* Water Rates from the [Open-Water-Rates-Specification](https://github.com/California-Data-Collaborative/Open-Water-Rate-Specification).
* Water Consumption data reported by the water agencies [ADD MORE DETAIL] 
* Demographic Data from the American Community Survey [??]
* Qualitative Data from a Survey realized by the California Data Collaborative with water agencies in 2017. 

```{r, include=FALSE}
source("R/utils.R")
source("R/plots.R")

#starting function to create Figure captions
fig_nums <- captioner(prefix = "Figure")


#Declare the customer classes to be tested for each utility
customer_classes <- c("RESIDENTIAL_SINGLE"
                      #"RESIDENTIAL_MULTI",
                      #"COMMERCIAL"
)
#End

#import sample data and give the data parameters used in the owrs files that are not in the original data
# df_smc <- santamonica
# names(df_smc) <- c( "cust_id", "usage_ccf", "usage_month", "usage_year",  "cust_class", "usage_date", "cust_class_from-utility")

# df_sample <- tbl_df(df_smc) %>%
#   mutate(hhsize = 4, meter_size = '3/4"', usage_zone = 1, landscape_area = 2000,
#          et_amount = 2.0, wrap_customer = "No", irr_area = 2000, carw_customer = "No",
#          season = "Winter", tax_exemption = "granted", lot_size_group = 3,
#          temperature_zone = "Medium", pressure_zone = 1, water_font = "city_delivered",
#          area = "inside_city", water_type = "potable", rate_class = "C1",
#          dwelling_units = 10, elevation_zone = 2, greater_than = "False",
#          usage_indoor_budget_ccf = .3, meter_type = "compound", block = 1,
#          tariff_area = 1, turbine_meter = "No", senior = "no") %>%
#   group_by(cust_class)

df_adjustable_sample <- tbl_df(data.frame (usage_month = 7, days_in_period = 31, usage_year=2017,
                                           hhsize = 4, meter_size = '3/4"', usage_zone = 1, landscape_area = 2000,
                                           et_amount = 7.0, wrap_customer = "No", irr_area = 2000, 
                                           carw_customer = "No", season = "Summer", tax_exemption = "granted", 
                                           lot_size_group = 3, temperature_zone = "Medium", pressure_zone = 1, 
                                           water_font = "city_delivered", city_limits = "inside_city", 
                                           water_type = "potable", rate_class = "C1", dwelling_units = 10, 
                                           elevation_zone = 2, greater_than = "False", usage_indoor_budget_ccf = .3, 
                                           meter_type = "Turbine", block = 1, tariff_area = 1, turbine_meter = "No",
                                           senior = "no", cust_class = customer_classes[1]))

#change for The scatter plots (If start = end the process will be a lot faster if only concerned with histograms and bar/pie charts)
start <- 0;
end <- 50;
interval <- 5;
df_usage <- as.data.frame(list("usage_ccf"=15, "cust_class"="RESIDENTIAL_SINGLE"))
df_sample <-  left_join(df_usage, df_adjustable_sample, by="cust_class")



#Retrieve the directories and files in the directores from the Open-Water-Specification-File directory
owrs_path <- "../Open-Water-Rate-Specification/full_utility_rates/California";

#TODO insert the filename gathering functioon here
df_OWRS <- tbl_df(as.data.frame(list("filepath"=getFileNames(owrs_path)), stringsAsFactors=FALSE)) %>% 
  
  mutate(owrs_directory = map(filepath, strsplit, split="/") %>% map(c(1,1))) %>%
  
  mutate(filename = map(filepath, strsplit, split="/") %>% map(c(1,2))) %>%
  
  mutate(utility_id = map(owrs_directory, strsplit, split=" ") %>%  
                      map(1) %>%map(tail, n=1) %>%
                      map(gsub, pattern="\\D", replacement="") %>%
                      map(as.numeric)) %>%
  
  mutate(effective_date = sapply(filename, extract_date) ) %>% 
  mutate(utility_name = sapply(as.character(owrs_directory), extract_utility_name) )


df_bill <- calculate_bills_for_all_utilities(df_OWRS, df_sample, owrs_path, customer_classes)

#Format the Bill Information so that only valid data entries are presented, the decimal points are rounded, and the data is arranged by utility
df_final_bill <- tbl_df(df_bill) %>% filter(!is.na(bill)) %>%
  mutate(bill = round(as.numeric(bill), 2),
         commodity_charge = round(as.numeric(commodity_charge), 2),
         service_charge = round(service_charge, 2),
         utility_name = as.character(utility_name),
         bill_frequency = as.character(bill_frequency),
         usage_ccf = ifelse(unit_type=="kgal", 1.33681*usage_ccf, usage_ccf)) %>% 
  arrange(utility_name)
#End

#getting rid of the annually charged bills until we fix the bill calculator
df_final_bill <- df_final_bill[which(df_final_bill$bill_frequency != "Annually"),]
```

```{r, echo=FALSE}

# #generating csv with rates for merge with demographics
# 
# #getting rid of the more recent years
# df_final_bill$effective_date <- as.Date(df_final_bill$effective_date)        #convert effective_date to Date
# df_final_bill <- df_final_bill[year(df_final_bill$effective_date) < 2018,]   # drop years >= 2018
# 
# #getting rid of duplicates
# #making sure to keep the most recent values
# df_final_bill <- df_final_bill[rev(order(df_final_bill$utility_name, df_final_bill$effective_date)),]
# #drop duplicates
# df_final_bill <- df_final_bill[ !duplicated(df_final_bill$utility_name), ]
# df_final_bill <- df_final_bill[order(df_final_bill$utility_name),] #reorder alphabetically
# 
# #adding pwsid
# supplier_pwsid <- read.csv('data/utilities_for_OWRS.csv', stringsAsFactors=FALSE)
# df_final_bill$fuzzy_match <- as.character(sapply(df_final_bill$utility_name, GetCloseMatches,
#                               sequence_strings = supplier_pwsid$Agency_Name, n=1L, cutoff = 0.85))
# df_final_bill <- merge(df_final_bill, supplier_pwsid, by.x = "fuzzy_match", by.y = "Agency_Name", all.x=TRUE, all.y=FALSE)
# 
# #write csv with the df_final_bill data frame
# df_final_bill$utility_id <- as.character(df_final_bill$utility_id)
# write.csv(df_final_bill, "data/bill_15ccf_residential.csv")

```


```{r, echo=FALSE, message=FALSE}
#library(rgdal)

#shape <- readOGR(dsn = "../Shapefiles", layer = "service_areas_cadc_with_utility_id")
#shape <- merge(shape, df_final_bill, by = "utility_id")

#setwd("../../Dropbox/CSV/")
#Output The Final Bill information to a  shapefile
#writeOGR(shape, ".", "service_areas_cadc_with_billing_info", driver="ESRI Shapefile", overwrite_layer = TRUE)
#End
#zipDemFiles <- list.files(recursive = TRUE)
#setwd("../")
#zip(zipfile = "../Dropbox/CSV/shapefileZip", files = paste("TempShapeFiles", zipDemFiles, sep = "/"))

#setwd("../../Documents/WaterRateTester/")
```

# Summary Statistics

This section discusses general characteristics of the rates for utilities analyzed in this survey.

```{r, echo=FALSE}
#The usage_ccf to evaluate for the histograms and bar chart
singleTargetValue <- 10;

```

```{r bill_frequency_pie, echo=FALSE, fig.cap = bill_freq_cap}
plot_bill_frequency_piechart(df_final_bill)

bill_freq_cap <- fig_nums(name='bill_frequency_piechart', caption= 'Bill Frequency Pie Chart. About three quarters of the water agencies use a monthly billing system.')
```

```{r mean_bill_pie, echo=FALSE, fig.cap = mean_bill_pie_cap}
plot_mean_bill_pie(df_final_bill, singleTargetValue)

mean_bill_pie_cap <- fig_nums(name='mean_bill_pie', caption= 'Average bill by parts for all agencies, considering a consumption of 10 CCF in a month. The average total bill is $60.68. With an average service charge (fixed) of $24.63 (40.6%) and an average commodity charge (variable) of $35.61 (58.7%).')
```



```{r rate_type_pie, echo=FALSE}
plot_rate_type_pie(df_final_bill)


```

```{r commodity_charge_vs_usage, echo=FALSE}

plot_commodity_charges_vs_usage(df_final_bill, start, end, interval)
```

```{r, echo=FALSE}
# plot_bills_vs_usage(df_final_bill, start, end, interval)
```

```{r, echo=FALSE}

# png(filename='plots/boxplot_bills.png',  width = 700, height = 432, units = "px")
# plot(boxplot_bills_vs_usage(df_final_bill, start, end, interval))
# dev.off()

boxplot_bills_vs_usage(df_final_bill, start, end, interval)
```




```{r, echo=FALSE}
# meanpercentFixed <- round(mean(as.numeric(df_final_bill$percentFixed[df_final_bill$usage_ccf == singleTargetValue])), 3)


plot_ratio_histogram(df_final_bill, singleTargetValue)
```

```{r, echo=FALSE}
plot_bill_histogram(df_final_bill, singleTargetValue)
```

# Rates x Efficiency
## Define Period of Analysis
```{r, echo=FALSE}
# startmonth <- 1
# startyear <- 2017
# endmonth <- 12
# endyear <- 2017
# 
# startdate <- as.Date(paste('01', as.character(startmonth), as.character(startyear)), "%d %m %Y")
# enddate <- as.Date(paste('01', as.character(endmonth), as.character(endyear)), "%d %m %Y")
# reference_period <- seq.Date(startdate, enddate, by="month")

```
## Calculate Rates
```{r, echo=FALSE}
# Rates time series

df_usage <- as.data.frame(list("usage_ccf"=15, "cust_class"="RESIDENTIAL_SINGLE"))

# df_adjustable_sample <- tbl_df(data.frame (reference_date = reference_period,
#                                            usage_month = month(reference_period),
#                                            usage_year = year(reference_period),
#                                            days_in_period = days_in_month(reference_period),
#                                            hhsize = 4, meter_size = '3/4"', usage_zone = 1, landscape_area = 2000,
#                                            et_amount = 7.0, wrap_customer = "No", irr_area = 2000,
#                                            carw_customer = "No", season = "Winter", tax_exemption = "granted",
#                                            lot_size_group = 3, temperature_zone = "Medium", pressure_zone = 1,
#                                            water_font = "city_delivered", city_limits = "inside_city",
#                                            water_type = "potable", rate_class = "C1", dwelling_units = 10,
#                                            elevation_zone = 2, greater_than = "False", usage_indoor_budget_ccf = .3,
#                                            meter_type = "Turbine", block = 1, tariff_area = 1, turbine_meter = "No",
#                                            senior = "no", cust_class = customer_classes[1]))

df_adjustable_sample$reference_date <- as.Date("2017-06-01")

df_sample <-  left_join(df_usage, df_adjustable_sample, by="cust_class")

df_bill <- calculate_bills_for_all_utilities(df_OWRS, df_sample, owrs_path, customer_classes)

#list of utilities for which the 3/4" key was not found to calculate the rates
#identified from the errors thrown when running this chunk of code.
meter_size_58 <- c('Goleta Water District',
                   'Humboldt Bay Municipal Water District', 
                   'Mid-Peninsula Water District', 
                   'North Marin Water District')

df_sample$meter_size <- '5/8"'

df_bill <- rbind(df_bill,
                 calculate_bills_for_all_utilities(df_OWRS[df_OWRS$utility_name %in% meter_size_58,], 
                                                   df_sample, owrs_path, customer_classes))
#same process for the utilities where the smallest meter size is 1 inch.
meter_size_1in <- c('Humboldt Community Services District')

df_sample$meter_size <- '1"'

df_bill <- rbind(df_bill,
                 calculate_bills_for_all_utilities(df_OWRS[df_OWRS$utility_name %in% meter_size_1in,], 
                                                   df_sample, owrs_path, customer_classes))

#Format the Bill Information so that only valid data entries are presented, 
#the decimal points are rounded, and the data is arranged by utility
df_final_bill <- tbl_df(df_bill) %>% filter(!is.na(bill)) %>%
  mutate(bill = round(as.numeric(bill), 2),
         commodity_charge = round(as.numeric(commodity_charge), 2),
         service_charge = round(service_charge, 2),
         utility_name = as.character(utility_name),
         bill_frequency = as.character(bill_frequency),
         usage_ccf = ifelse(unit_type=="kgal", 1.33681*usage_ccf, usage_ccf)) %>%
  filter(as.Date(reference_date) >= as.Date(effective_date) ) %>% 
  group_by(utility_name, reference_date) %>%
  filter(as.Date(effective_date) == max(as.Date(effective_date) )) %>%
  arrange(utility_name)

#getting rid of the annually charged bills until we fix the bill calculator
df_final_bill <- df_final_bill[which(df_final_bill$bill_frequency != "Annually"),]
```
Average water rates history:
```{r, echo=FALSE}
# plot_avg_price_history(df_final_bill)
```


## Calculate Efficiency
Load suppliers report info and join with the Utilities list from the OWRS files
```{r, echo=FALSE}
#load supplier reports, geoinformation and pwsid_record
supplier_reports <- read.csv('data/supplier_report.csv', stringsAsFactors=FALSE)
#supplier_geo <- read.csv('data/suppliers.csv', stringsAsFactors=FALSE)
supplier_pwsid <- read.csv('data/utilities_for_OWRS.csv', stringsAsFactors=FALSE)



# append to df_OWRS the best fuzzy match for utility_name to get pwsid
# cutoff chosen arbitraily, other values can be tested
df_OWRS$fuzzy_match <- as.character(sapply(df_OWRS$utility_name, GetCloseMatches,
                              sequence_strings = supplier_pwsid$Agency_Name, n=1L, cutoff = 0.85))

# get pswid
merged_OWRS <- merge(df_OWRS, supplier_pwsid, by.x = "fuzzy_match", by.y = "Agency_Name", all.x=TRUE, all.y=FALSE)
# merge with suplier report
merged_OWRS <- merge(merged_OWRS, supplier_reports, by.x = "PWSID", by.y = "report_pwsid", all.x=TRUE, all.y=FALSE)
```
Calculate Efficiency from the suppliers reports
```{r, echo=FALSE}
#The standard value for GPCD is assumed 55
target_gpcd <- 55
ET_adj_factor <- 0.8
unit_conversion <- 0.62

merged_OWRS$production_target <- 55 * merged_OWRS$report_population * merged_OWRS$report_days_in_month +
  merged_OWRS$report_irr_area_sf * merged_OWRS$report_eto * ET_adj_factor * unit_conversion

merged_OWRS$residential_use <- merged_OWRS$report_production_calculated * merged_OWRS$report_percent_residential

merged_OWRS$pct_above_target <- (merged_OWRS$residential_use / merged_OWRS$production_target) - 1

merged_OWRS$report_monthyear <- as.Date(paste('01', as.character(merged_OWRS$report_month),
                                              as.character(merged_OWRS$report_year)), "%d %m %Y")
```

```{r, echo=FALSE}
plot_efficiency_ts(merged_OWRS)

```

```{r, echo=FALSE}
plot_gpcd_ts(merged_OWRS)
```
## Compare Rates and efficiency
```{r, echo=FALSE}
df_final_bill$fuzzy_match <- as.character(sapply(df_final_bill$utility_name, GetCloseMatches,
                              sequence_strings = supplier_pwsid$Agency_Name, n=1L, cutoff = 0.85))

df_final_bill <- merge(df_final_bill, supplier_pwsid, by.x = "fuzzy_match", by.y = "Agency_Name", all.x=TRUE, all.y=FALSE)


#merge on PWSID, month and Year
eff_vs_rate <- merge(df_final_bill, merged_OWRS, 
                     by.x = c("PWSID", "usage_month", "usage_year"),
                     by.y = c("PWSID", "report_month", "report_year"),
                     all.x=TRUE, all.y=FALSE)
```

Scatter plot of Efficiency (pct_above_target) vs Rates (Total Bill for 15 CCF)
```{r, echo=FALSE}
agg_eff_vs_rate <- eff_vs_rate[c("PWSID", "usage_month", "usage_year",
                                 "bill", "pct_above_target")] %>% na.omit() %>%
                        group_by(PWSID) %>% summarise_all(funs(mean))

plot_eff_vs_bill(agg_eff_vs_rate)
```

Scatter plot of Efficiency vs Rates Structure (% Fixed  - for 15 CCF)
```{r, echo=FALSE}

agg_eff_vs_rate <- eff_vs_rate[c("PWSID", "usage_month", "usage_year",
                                 "pct_above_target", "percentFixed")] %>% na.omit() %>%
                        group_by(PWSID) %>% summarise_all(funs(mean))

plot_eff_vs_pctFixed(agg_eff_vs_rate)
```

## Joining Data from the Qualitative Survey
```{r, echo=FALSE}
#read the survey data in
hd <- read.csv('data/2017 CA-NV Rate Survey.csv', nrows=2, header=FALSE, stringsAsFactors = FALSE)
quali_survey <- read.csv('data/2017 CA-NV Rate Survey.csv', skip=2, header=FALSE, stringsAsFactors = FALSE)
names(quali_survey) <- paste(hd[1,], hd[2,], sep = "#")
names(quali_survey)[36] <- "costs_pct_fixed"
names(quali_survey)[37] <- "rev_pct_fixed"
names(quali_survey)[10] <- "agency_name"
names(quali_survey)[5] <- "IP_address"
names(quali_survey)[14] <- "city-town"

quali_survey$costs_pct_fixed <- as.numeric(quali_survey$costs_pct_fixed)
quali_survey$rev_pct_fixed <- as.numeric(quali_survey$rev_pct_fixed)

```
```{r}
pct_costs <- ggplot(quali_survey, aes(costs_pct_fixed)) + geom_histogram(bins = 10)

pct_costs
```



```{r}
pct_rev <- ggplot(quali_survey, aes(rev_pct_fixed)) + geom_histogram(bins = 10)

pct_rev
```

```{r}
quali_survey$fixedRev_per_fixedCosts <- quali_survey$rev_pct_fixed / quali_survey$costs_pct_fixed

ggplot(quali_survey, aes(fixedRev_per_fixedCosts)) + geom_histogram(binwidth = 0.5)
```
```{r}



```

