---
title: "California Water Rate Survey Results 2017"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(scales)
library(raster)
library(RateParser)
library(yaml)
library(purrr)
library(fuzzywuzzyR)
library(lubridate)
library(captioner)
```

# Introduction

The [California Data Collaborative ("CaDC")](http://californiadatacollaborative.org/) is a coalition of water utilities that have pioneered a new data infrastructure non-profit 501 (c) (3) to support water managers in meeting their reliability objectives and serve the public good.

One important contribution of the CaDC was to establish a standard format and to provide the infrastructure for storage and maintainance of an open database for water rates, facilitating the work of analysts, economists and software developers interested in analyzing and understanding the differences in water rate structures and prices across many different agencies and locations. The water rate structures were organized in [Open-Water-Rate-Specification (OWRS)](https://github.com/California-Data-Collaborative/Open-Water-Rate-Specification) files, a format based on [YAML](http://yaml.org/), which is designed to be easy to store, transmit, and parse in any programming language while also being easy for humans to read.

This report presents a summary of the types of analyses and insights that can be obtained from analyzing the OWRS, especially when this information is combined with the water consumption data from water agencies and Census Data.

# Data

This report provides the combined analysis of data from 4 different sources:

* Water Rates from the [Open-Water-Rates-Specification](https://github.com/California-Data-Collaborative/Open-Water-Rate-Specification).
* Water Consumption data reported by the water agencies [ADD MORE DETAIL] 
* Demographic Data from the American Community Survey [??]
* Qualitative Data from a Survey realized by the California Data Collaborative with water agencies in 2017. 

```{r, include=FALSE}
source("R/utils.R")
source("R/plots.R")

#starting function to create Figure captions
fig_nums <- captioner(prefix = "Figure")


#Declare the customer classes to be tested for each utility
customer_classes <- c("RESIDENTIAL_SINGLE"
                      #"RESIDENTIAL_MULTI",
                      #"COMMERCIAL"
)
#End

#import sample data and give the data parameters used in the owrs files that are not in the original data
df_smc <- santamonica
names(df_smc) <- c( "cust_id", "usage_ccf", "usage_month", "usage_year",  "cust_class", "usage_date", "cust_class_from-utility")

df_sample <- tbl_df(df_smc) %>%
  mutate(hhsize = 4, meter_size = '3/4"', usage_zone = 1, landscape_area = 2000,
         et_amount = 2.0, wrap_customer = "No", irr_area = 2000, carw_customer = "No",
         season = "Winter", tax_exemption = "granted", lot_size_group = 3,
         temperature_zone = "Medium", pressure_zone = 1, water_font = "city_delivered",
         area = "inside_city", water_type = "potable", rate_class = "C1",
         dwelling_units = 10, elevation_zone = 2, greater_than = "False",
         usage_indoor_budget_ccf = .3, meter_type = "compound", block = 1,
         tariff_area = 1, turbine_meter = "No", senior = "no") %>%
  group_by(cust_class)

df_adjustable_sample <- tbl_df(data.frame (usage_month = 7, days_in_period = 31, usage_year=2017,
                                           hhsize = 4, meter_size = '3/4"', usage_zone = 1, landscape_area = 2000,
                                           et_amount = 7.0, wrap_customer = "No", irr_area = 2000, 
                                           carw_customer = "No", season = "Summer", tax_exemption = "granted", 
                                           lot_size_group = 3, temperature_zone = "Medium", pressure_zone = 1, 
                                           water_font = "city_delivered", city_limits = "inside_city", 
                                           water_type = "potable", rate_class = "C1", dwelling_units = 10, 
                                           elevation_zone = 2, greater_than = "False", usage_indoor_budget_ccf = .3, 
                                           meter_type = "Turbine", block = 1, tariff_area = 1, turbine_meter = "No",
                                           senior = "no", cust_class = customer_classes[1]))

#change for The scatter plots (If start = end the process will be a lot faster if only concerned with histograms and bar/pie charts)
start <- 0;
end <- 30;
interval <- 5;
df_usage <- as.data.frame(list("usage_ccf"= seq(from = start, to = end, by = interval),
                               "cust_class"="RESIDENTIAL_SINGLE"))
df_sample <-  left_join(df_usage, df_adjustable_sample, by="cust_class")



#Retrieve the directories and files in the directores from the Open-Water-Specification-File directory
owrs_path <- "../Open-Water-Rate-Specification/full_utility_rates/California";

#TODO insert the filename gathering functioon here
df_OWRS <- tbl_df(as.data.frame(list("filepath"=getFileNames(owrs_path)), stringsAsFactors=FALSE)) %>% 
  
  mutate(owrs_directory = map(filepath, strsplit, split="/") %>% map(c(1,1))) %>%
  
  mutate(filename = map(filepath, strsplit, split="/") %>% map(c(1,2))) %>%
  
  mutate(utility_id = map(owrs_directory, strsplit, split=" ") %>%  
                      map(1) %>%map(tail, n=1) %>%
                      map(gsub, pattern="\\D", replacement="") %>%
                      map(as.numeric)) %>%
  
  mutate(effective_date = sapply(filename, extract_date) ) %>% 
  mutate(utility_name = sapply(as.character(owrs_directory), extract_utility_name) )


df_bill <- calculate_bills_for_all_utilities(df_OWRS, df_sample, owrs_path, customer_classes)

#list of utilities for which the 3/4" key was not found to calculate the rates
#identified from the errors thrown when running this chunk of code.
meter_size_58 <- c('Goleta Water District',
                   'Humboldt Bay Municipal Water District', 
                   'Mid-Peninsula Water District', 
                   'North Marin Water District')

df_sample$meter_size <- '5/8"'

df_bill <- rbind(df_bill,
                 calculate_bills_for_all_utilities(df_OWRS[df_OWRS$utility_name %in% meter_size_58,], 
                                                   df_sample, owrs_path, customer_classes))
#same process for the utilities where the smallest meter size is 1 inch.
meter_size_1in <- c('Humboldt Community Services District')

df_sample$meter_size <- '1"'

df_bill <- rbind(df_bill,
                 calculate_bills_for_all_utilities(df_OWRS[df_OWRS$utility_name %in% meter_size_1in,], 
                                                   df_sample, owrs_path, customer_classes))

#Format the Bill Information so that only valid data entries are presented, the decimal points are rounded, and the data is arranged by utility
df_final_bill <- tbl_df(df_bill) %>% filter(!is.na(bill)) %>%
  mutate(bill = round(as.numeric(bill), 2),
         commodity_charge = round(as.numeric(commodity_charge), 2),
         service_charge = round(service_charge, 2),
         utility_name = as.character(utility_name),
         bill_frequency = as.character(bill_frequency),
         usage_ccf = ifelse(unit_type=="kgal", 1.33681*usage_ccf, usage_ccf)) %>% 
  arrange(utility_name)
#End

#getting rid of the annually charged bills until we fix the bill calculator
df_final_bill <- df_final_bill[which(df_final_bill$bill_frequency != "Annually"),]
```

```{r, echo=FALSE}

# #generating csv with rates for merge with demographics
# 
# #getting rid of the more recent years
# df_final_bill$effective_date <- as.Date(df_final_bill$effective_date)        #convert effective_date to Date
# df_final_bill <- df_final_bill[year(df_final_bill$effective_date) < 2018,]   # drop years >= 2018
# 
# #getting rid of duplicates
# #making sure to keep the most recent values
# df_final_bill <- df_final_bill[rev(order(df_final_bill$utility_name, df_final_bill$effective_date)),]
# #drop duplicates
# df_final_bill <- df_final_bill[ !duplicated(df_final_bill$utility_name), ]
# df_final_bill <- df_final_bill[order(df_final_bill$utility_name),] #reorder alphabetically
# 
# #adding pwsid
# supplier_pwsid <- read.csv('data/utilities_for_OWRS.csv', stringsAsFactors=FALSE)
# df_final_bill$fuzzy_match <- as.character(sapply(df_final_bill$utility_name, GetCloseMatches,
#                               sequence_strings = supplier_pwsid$Agency_Name, n=1L, cutoff = 0.85))
# df_final_bill <- merge(df_final_bill, supplier_pwsid, by.x = "fuzzy_match", by.y = "Agency_Name", all.x=TRUE, all.y=FALSE)
# 
# #write csv with the df_final_bill data frame
# df_final_bill$utility_id <- as.character(df_final_bill$utility_id)
# write.csv(df_final_bill, "data/bill_15ccf_residential.csv")

```


```{r, echo=FALSE, message=FALSE}
#library(rgdal)

#shape <- readOGR(dsn = "../Shapefiles", layer = "service_areas_cadc_with_utility_id")
#shape <- merge(shape, df_final_bill, by = "utility_id")

#setwd("../../Dropbox/CSV/")
#Output The Final Bill information to a  shapefile
#writeOGR(shape, ".", "service_areas_cadc_with_billing_info", driver="ESRI Shapefile", overwrite_layer = TRUE)
#End
#zipDemFiles <- list.files(recursive = TRUE)
#setwd("../")
#zip(zipfile = "../Dropbox/CSV/shapefileZip", files = paste("TempShapeFiles", zipDemFiles, sep = "/"))

#setwd("../../Documents/WaterRateTester/")
```

# Rate Structures

This section discusses general characteristics of the rates for utilities analyzed in this survey.

```{r, echo=FALSE}
#The usage_ccf to evaluate for the histograms and bar chart
singleTargetValue <- 15;

```

```{r bill_frequency_pie, echo=FALSE, fig.cap = bill_freq_cap}

bill_frequency_pie <- plot_bill_frequency_piechart(df_final_bill %>% filter(usage_ccf==15))

bill_frequency_pie
bill_freq_cap <- fig_nums(name='bill_frequency_piechart', caption= 'Bill Frequency Pie Chart.')
```

`r fig_nums("bill_frequency_piechart", display = "cite")` above shows that `r round(bill_frequency_pie$data[1, 'Value'] / sum(bill_frequency_pie$data[, 'Value']) * 100, digits=1)`% of the agencies charge their customers on a monthly basis.
That indicates an increase in relation to the previous survey (2015), showing that there seems to be trend in Agencies moving from a bi-monthly billing system to a monthly one. 

```{r mean_bill_pie, echo=FALSE, fig.cap = mean_bill_pie_cap}
mean_bill_pie <- plot_mean_bill_pie(df_final_bill, singleTargetValue)
mean_bill_pie
mean_bill_pie_cap <- fig_nums(name='mean_bill_pie', caption= 'Average bill by parts for all agencies, considering a consumption of 15 CCF in a month.')
```

In terms of the different parts from a bill, `r fig_nums("mean_bill_pie", display = "cite")` shows that for a usage of `r singleTargetValue` CCF the commodity charge portion from the bill is the biggest one representing on average `r round(mean_bill_pie$data[2, 'Value'] / sum(bill_frequency_pie$data[, 'Value']) * 100, digits=1)`% of the total bill.

```{r rate_type_pie, echo=FALSE, fig.cap = rate_type_pie_cap}
rate_type_pie <- plot_rate_type_pie(df_final_bill)
rate_type_pie
rate_type_pie_cap <- fig_nums(name='rate_type_pie', caption= 'Types of rate structures.')

```

From analyzing the different rate structures adopted by each agency, `r fig_nums("rate_type_pie", display = "cite")` shows that most of the agencies(`r round(rate_type_pie$data[2, 'Value'] / sum(rate_type_pie$data[, 'Value']) * 100, digits=1)`%) adopt a tiered rate structure, but a significant amount still adopt an uniform rate structure (`r round(rate_type_pie$data[3, 'Value'] / sum(rate_type_pie$data[, 'Value']) * 100, digits=1)`%)

```{r commodity_charge_vs_usage, echo=FALSE}

# commodity_charge_vs_usage <- plot_commodity_charges_vs_usage(df_final_bill, start, end, interval)
# commodity_charge_vs_usage
# comm_charge_vs_usage_cap <- fig_nums(name='commodity_charge_vs_usage', caption= 'Commodity Charge vs Usage.')
```


```{r total_bill_vs_usage, echo=FALSE }
# bills_vs_usage <- plot_bills_vs_usage(df_final_bill, start, end, interval)
# bills_vs_usage
# total_bill_vs_usage <- fig_nums(name='total_bill_vs_usage', caption= 'Total Bill vs Usage.')
```


```{r service_ratio_histogram, echo=FALSE, fig.cap = ratio_histogram_cap}

service_ratio_histogram <- plot_ratio_histogram(df_final_bill, singleTargetValue)
service_ratio_histogram
ratio_histogram_cap <- fig_nums(name='service_ratio_histogram', caption= 'Ratio of Service charge to total bill at 15 CCF.')
```

The histogram in `r fig_nums("service_ratio_histogram", display = "cite")` shows the distribution and the mean of the ratio between the service charge and the total bill for each agency. That number is a good proxy for the proportion of revenue in each agency that is fixed and does not depend on the consumer's water usage.

```{r bill_histogram, echo=FALSE, fig.cap = bill_histogram_cap}

bill_histogram <- plot_bill_histogram(df_final_bill, singleTargetValue)
bill_histogram
bill_histogram_cap <- fig_nums(name='bill_histogram', caption= 'Total bill at 15 CCF.')
```

`r fig_nums("bill_histogram", display = "cite")` shows an histogram of the different amounts charged by each agency for a water usage of 15 CCF in single family residential units.

```{r total_bill_vs_usage_box, echo=FALSE, warning=FALSE, fig.cap = total_bill_vs_usage_box_cap}

total_bill_vs_usage_box <- boxplot_bills_vs_usage(df_final_bill, start, end, interval)
total_bill_vs_usage_box
total_bill_vs_usage_box_cap <- fig_nums(name='total_bill_vs_usage_box', caption= 'Boxplot of Total Bill vs Usage.')
```

`r fig_nums("total_bill_vs_usage_box", display = "cite")` ilustrates how the variability of the total water bill among agencies increases the higher the water usage by single residential units. It noteworthy that at a usage as low as 10 to 15 CCF there can be already a difference of two to three fold between agencies. 

# Relating Rates to Efficiency

This section provides an analysis of the relationships between the rates charged by each water agency and the efficiency in water use in the areas served by those agencies.

```{r, echo=FALSE}
startmonth <- 09
startyear <- 2016
endmonth <- 08
endyear <- 2017

startdate <- as.Date(paste('01', as.character(startmonth), as.character(startyear)), "%d %m %Y")
enddate <- as.Date(paste('01', as.character(endmonth), as.character(endyear)), "%d %m %Y")
reference_period <- seq.Date(startdate, enddate, by="month")


```
The period considered for the analysis is `r startmonth`/`r startyear` to `r endmonth`/`r endyear`.


## Rates in the period
```{r, include=FALSE}
# Rates time series

df_usage <- as.data.frame(list("usage_ccf"=15, "cust_class"="RESIDENTIAL_SINGLE"))

df_adjustable_sample <- tbl_df(data.frame (reference_date = reference_period,
                                           usage_month = month(reference_period),
                                           usage_year = year(reference_period),
                                           days_in_period = days_in_month(reference_period),
                                           hhsize = 4, meter_size = '3/4"', usage_zone = 1, landscape_area = 2000,
                                           et_amount = 7.0, wrap_customer = "No", irr_area = 2000,
                                           carw_customer = "No", season = "Winter", tax_exemption = "granted",
                                           lot_size_group = 3, temperature_zone = "Medium", pressure_zone = 1,
                                           water_font = "city_delivered", city_limits = "inside_city",
                                           water_type = "potable", rate_class = "C1", dwelling_units = 10,
                                           elevation_zone = 2, greater_than = "False", usage_indoor_budget_ccf = .3,
                                           meter_type = "Turbine", block = 1, tariff_area = 1, turbine_meter = "No",
                                           senior = "no", cust_class = customer_classes[1]))

#df_adjustable_sample$reference_date <- as.Date("2017-07-01")

df_sample <-  left_join(df_usage, df_adjustable_sample, by="cust_class")

df_bill <- calculate_bills_for_all_utilities(df_OWRS, df_sample, owrs_path, customer_classes)

#list of utilities for which the 3/4" key was not found to calculate the rates
#identified from the errors thrown when running this chunk of code.
meter_size_58 <- c('Goleta Water District',
                   'Humboldt Bay Municipal Water District', 
                   'Mid-Peninsula Water District', 
                   'North Marin Water District')

df_sample$meter_size <- '5/8"'

df_bill <- rbind(df_bill,
                 calculate_bills_for_all_utilities(df_OWRS[df_OWRS$utility_name %in% meter_size_58,], 
                                                   df_sample, owrs_path, customer_classes))
#same process for the utilities where the smallest meter size is 1 inch.
meter_size_1in <- c('Humboldt Community Services District')

df_sample$meter_size <- '1"'

df_bill <- rbind(df_bill,
                 calculate_bills_for_all_utilities(df_OWRS[df_OWRS$utility_name %in% meter_size_1in,], 
                                                   df_sample, owrs_path, customer_classes))

#Format the Bill Information so that only valid data entries are presented, 
#the decimal points are rounded, and the data is arranged by utility
df_final_bill <- tbl_df(df_bill) %>% filter(!is.na(bill)) %>%
  mutate(bill = round(as.numeric(bill), 2),
         commodity_charge = round(as.numeric(commodity_charge), 2),
         service_charge = round(service_charge, 2),
         utility_name = as.character(utility_name),
         bill_frequency = as.character(bill_frequency),
         usage_ccf = ifelse(unit_type=="kgal", 1.33681*usage_ccf, usage_ccf)) %>%
  filter(as.Date(reference_date) >= as.Date(effective_date) ) %>% 
  group_by(utility_name, reference_date) %>%
  filter(as.Date(effective_date) == max(as.Date(effective_date) )) %>%
  arrange(utility_name)

#getting rid of the annually charged bills until we fix the bill calculator
df_final_bill <- df_final_bill[which(df_final_bill$bill_frequency != "Annually"),]
```


```{r avg_price_history, echo=FALSE, warning=FALSE, fig.cap= avg_price_history_cap}

plot_avg_price_history(df_final_bill)
avg_price_history_cap <- fig_nums(name='avg_price_history', caption= 'Average price charged for 15 CCF across all agencies included in the survey.')

```

`r fig_nums("avg_price_history", display = "cite")` shows that the prices remained fairly stable throughout the year. Although there was a large variability across agencies the the total bill for 15 CCF of usage for about half of the agencies stayed in the interval between $50 and $80 for most of the year. 

## Efficiency in the period
Load suppliers report info and join with the Utilities list from the OWRS files
```{r, echo=FALSE}
#load supplier reports, geoinformation and pwsid_record
supplier_reports <- read.csv('data/supplier_report.csv', stringsAsFactors=FALSE)
#supplier_geo <- read.csv('data/suppliers.csv', stringsAsFactors=FALSE)
supplier_pwsid <- read.csv('data/utilities_for_OWRS.csv', stringsAsFactors=FALSE)



# append to df_OWRS the best fuzzy match for utility_name to get pwsid
# cutoff chosen arbitraily, other values can be tested
df_OWRS$fuzzy_match <- as.character(sapply(df_OWRS$utility_name, GetCloseMatches,
                              sequence_strings = supplier_pwsid$Agency_Name, n=1L, cutoff = 0.85))

# get pswid
merged_OWRS <- merge(df_OWRS, supplier_pwsid, by.x = "fuzzy_match", by.y = "Agency_Name", all.x=TRUE, all.y=FALSE)
# merge with suplier report
merged_OWRS <- merge(merged_OWRS, supplier_reports, by.x = "PWSID", by.y = "report_pwsid", all.x=TRUE, all.y=FALSE)
```

```{r, echo=FALSE}
#The standard value for GPCD is assumed 55
target_gpcd <- 55
ET_adj_factor <- 0.8
unit_conversion <- 0.62

merged_OWRS$production_target <- 55 * merged_OWRS$report_population * merged_OWRS$report_days_in_month +
  merged_OWRS$report_irr_area_sf * merged_OWRS$report_eto * ET_adj_factor * unit_conversion

merged_OWRS$residential_use <- merged_OWRS$report_production_calculated * merged_OWRS$report_percent_residential

merged_OWRS$pct_above_target <- (merged_OWRS$residential_use / merged_OWRS$production_target) - 1

merged_OWRS$report_monthyear <- as.Date(paste('01', as.character(merged_OWRS$report_month),
                                              as.character(merged_OWRS$report_year)), "%d %m %Y")
```

```{r efficiency_ts, echo=FALSE, warning=FALSE, fig.cap= efficiency_ts_cap}
plot_efficiency_ts(merged_OWRS %>% filter(report_monthyear %in% reference_period))
efficiency_ts_cap <- fig_nums(name='efficiency_ts', caption= 'Efficiency measured as percentage above target.')

```

In terms of Efficiency `r fig_nums("efficiency_ts", display = "cite")` shows a considerable seasonal variation and that even with that variation, most agencies have met theri efficiency goals throughout the whole period. 

```{r, echo=FALSE}
#plot_gpcd_ts(merged_OWRS)
```

## Rates vs Efficiency
```{r, echo=FALSE}
df_final_bill$fuzzy_match <- as.character(sapply(df_final_bill$utility_name, GetCloseMatches,
                              sequence_strings = supplier_pwsid$Agency_Name, n=1L, cutoff = 0.85))

df_final_bill <- merge(df_final_bill, supplier_pwsid, by.x = "fuzzy_match", by.y = "Agency_Name", all.x=TRUE, all.y=FALSE)


#merge on PWSID, month and Year
eff_vs_rate <- merge(df_final_bill, merged_OWRS, 
                     by.x = c("PWSID", "usage_month", "usage_year"),
                     by.y = c("PWSID", "report_month", "report_year"),
                     all.x=TRUE, all.y=FALSE)
```


```{r eff_vs_total_bill, echo=FALSE, fig.cap = eff_vs_total_bill_cap}
agg_eff_vs_rate <- eff_vs_rate[c("PWSID", "usage_month", "usage_year",
                                 "bill", "pct_above_target")] %>% na.omit() %>%
                        group_by(PWSID) %>% summarise_all(funs(mean))

eff_vs_total_bill <- plot_eff_vs_bill(agg_eff_vs_rate)
eff_vs_total_bill
eff_vs_total_bill_cap <- fig_nums(name = 'eff_vs_total_bill', caption = 'Efficiency vs Total Bill.')
```

Despite the large variance around the linear fit, `r fig_nums("eff_vs_total_bill", display = "cite")` does show a significant[* pearson? another statistical test?] negative correlation between the percentage above target and the total bill paid by residential customers in single-family units. 

```{r eff_vs_pctFixed, echo=FALSE, fig.cap = eff_vs_pctFixed_cap}

agg_eff_vs_rate <- eff_vs_rate[c("PWSID", "usage_month", "usage_year",
                                 "pct_above_target", "percentFixed")] %>% na.omit() %>%
                        group_by(PWSID) %>% summarise_all(funs(mean))

eff_vs_pctFixed <- plot_eff_vs_pctFixed(agg_eff_vs_rate)
eff_vs_pctFixed
eff_vs_pctFixed_cap <- fig_nums(name = 'eff_vs_pctFixed', 
                                caption = 'Efficiency vs Percentage of Service Charge in Total Bill.')
```

Furthermore, `r fig_nums("eff_vs_pctFixed", display = "cite")` shows a positive correlation [significant? pearson?] of Percentage of Service Charge and Percentage Above Target. Indicating that the more sensitive their bill is to the amount of water they consume, more efficient the customers tend to be.  


## Operational Costs and Revenue in the Water Agencies
```{r, include=FALSE}
#read the survey data in
hd <- read.csv('data/2017 CA-NV Rate Survey.csv', nrows=2, header=FALSE, stringsAsFactors = FALSE)
quali_survey <- read.csv('data/2017 CA-NV Rate Survey.csv', skip=2, header=FALSE, stringsAsFactors = FALSE)
names(quali_survey) <- paste(hd[1,], hd[2,], sep = "#")
names(quali_survey)[36] <- "costs_pct_fixed"
names(quali_survey)[37] <- "rev_pct_fixed"
names(quali_survey)[10] <- "agency_name"
names(quali_survey)[5] <- "IP_address"
names(quali_survey)[14] <- "city-town"

quali_survey$costs_pct_fixed <- as.numeric(quali_survey$costs_pct_fixed)
quali_survey$rev_pct_fixed <- as.numeric(quali_survey$rev_pct_fixed)

```


```{r pct_fixed_costs, echo=FALSE, warning=FALSE, fig.cap=pct_fixed_costs_cap }
pct_fixed_costs <- ggplot(quali_survey, aes(costs_pct_fixed)) + 
                   geom_histogram(bins = 10, colour="black", fill="white") +
                   ggtitle('Percentage of Fixed Costs', 
                           subtitle = paste("Mean:", round(mean(quali_survey$costs_pct_fixed, na.rm=TRUE), 1),"%")) +
                   labs(y = "Number of Agencies", x = "Percentage of Fixed Costs") +
                   geom_vline(xintercept = mean(quali_survey$costs_pct_fixed, na.rm=TRUE), color = "red")

pct_fixed_costs
pct_fixed_costs_cap <- fig_nums(name = 'pct_fixed_costs', 
                                caption = 'Percentage of costs that are fixed for an Agency.')
```


```{r pct_fixed_rev, echo=FALSE, warning=FALSE, fig.cap=pct_fixed_rev_cap}
pct_fixed_rev <- ggplot(quali_survey, aes(rev_pct_fixed)) + 
                 geom_histogram(bins = 10, colour="black", fill="white")+
                 ggtitle('Percentage of Fixed Revenue', 
                         subtitle = paste("Mean:", round(mean(quali_survey$rev_pct_fixed, na.rm=TRUE), 1),"%")) +
                 labs(y = "Number of Agencies", x = " Percentage of Fixed Revenue") +
                 geom_vline(xintercept = mean(quali_survey$rev_pct_fixed, na.rm=TRUE), color = "red")

pct_fixed_rev
pct_fixed_rev_cap <- fig_nums(name='pct_fixed_rev', caption = "Percentage of an Agency's revenues that are fixed.")
```

In this year's survey, some water agencies indicated (as an optional response) how much of their revenue and how much of their costs come from fixed sources that do not depend on the amount of water usage in a given period. `r fig_nums("pct_fixed_costs", display = "cite")` shows that on average, the water agencies that provided that information, have `r round(mean(quali_survey$costs_pct_fixed, na.rm=TRUE))` % of their costs fixed, while `r round(mean(quali_survey$rev_pct_fixed, na.rm=TRUE))` % of their revenue being fixed, as show in `r fig_nums("pct_fixed_rev", display = "cite")`.


```{r fixedRev_per_fixedCosts, echo=FALSE, warning=FALSE, fig.cap= fixedRev_per_fixedCosts_cap}
quali_survey$fixedRev_per_fixedCosts <- quali_survey$rev_pct_fixed / quali_survey$costs_pct_fixed

fixedRev_per_fixedCosts <- ggplot(quali_survey, aes(fixedRev_per_fixedCosts)) + 
                           geom_histogram(binwidth = 0.2, colour="black", fill="white")+
                           ggtitle('Ratio of Fixed Revenue per Fixed Costs',
                                   subtitle = paste("Median:", round(median(quali_survey$fixedRev_per_fixedCosts, 
                                                                        na.rm=TRUE), 1))) +
                           labs(y = "Number of Agencies", x = " Ratio of % Fixed Revenue per  % Fixed Costs") +
                           geom_vline(xintercept = median(quali_survey$fixedRev_per_fixedCosts, 
                                                        na.rm=TRUE), color = "red")

fixedRev_per_fixedCosts
fixedRev_per_fixedCosts_cap <- fig_nums(name='fixedRev_per_fixedCosts', caption = "Ratio of Percentage of Fixed Revenue per Percentage of Fixed Costs.")
```

In addition, when computing the ratio between the percentage of fixed revenue to percentage of fixed costs, `r fig_nums("fixedRev_per_fixedCosts", display = "cite")` shows that at least half of the agencies have their percentage of fixed revenue smaller than or equal to the percentage of fixed costs.
